Cluster Difference Imaging Photometric Survey. IV. TOI 1937Ab. Youngest Hot Jupiter, or Just Tidally Spun Up?

### 2021-02-16 cluster_paper/ms.tex
update paper

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these halos could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$120\,Myr open cluster NGC\,2516.
  The classical tidal radius of the cluster is $\sim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as is the case with
  comparable stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

Star clusters form in overdensities of hierarchically structured
molecular clouds (CITE).  Over the first $\approx$10\,Myr,
``feedback'' effects including protostellar outflows, photoionization,
radiation pressure, and supernova shocks conspire to disperse the gas
out of the cloud (CITE Krumholz+19).  Since only a small fraction of
the cloud mass is converted into stars ($\approx 1\%$; see Krumholz+19
Figure 10), this gas dispersal drives the majority ($\sim$90\%) of the
stars in the cluster to become gravitationally unbound (e.g., Lada \&
Lada 03).

From 10 to 1000\,Myr, the portion of the cluster that survives gas
dispersal suffers an onslaught that almost always succeeds in
destroying the cluster.  Internal to the cluster, mass loss from
supernovae and AGB winds, and close stellar encounters
(``relaxation'') are important over 10s and 100s of megayears,
respectively.  External to the cluster, collisions with GMCs (Spitzer
1958), and perturbations from the galactic tide in both the radial
(CITE) and vertical (CITE) dimensions further drive stars toward the
``high velocity tails'' of the stellar energy distribution.

@@

Identification of the stars that disperse into the galactic field is an
important aspect of understanding the conditions under which stars and
star clusters form, and for understanding how they subsequently evolve.
For instance, how does the process of radial migration across the
galactic disk affect cluster dispersal?  Did the Sun form in an open
cluster? If so, how massive was its cluster, and is there any hope at
identifying the stars that formed near the Sun?  Qualitatively, the
dispersal of open clusters also provides perhaps the best test case for
the concept of ``chemical tagging'', aka.\ ``galactic archaeology''
(CITE).  If you can't chemically tag stars that are kinematically
associated, why would you be able to do so once they've dispersed
throughout the galactic disk?

Outside of the issue of star formation, identifying the remnant
halos\footnote{ The low-density outer regions of open clusters have
been referred to in general terms as ``halos'' and ``coronae''; we
adopt the former (CITE, CITE). } of open clusters is important for a
separate project: that of discovering young transiting planets.  Young
transiting planets are hard to find because young stars are rare
(e.g., CITE Piskunov 2018), and often reside in crowded regions of the
sky (e.g., CITE ANY OPEN CLUSTER CATALOG).  If the dispersed halos of
nearby star clusters could be reliably identified in a complete
manner, this could expand the census of nearby star clusters by up to
a factor of $10$, based on the expected fraction of stars thought to
be lost during gas dispersal.  Path to identifying these young stars
include photometric criteria (CITE Zari+18), spectroscopic criteria
(CITE Berger+18), and combinations thereof (e.g., CITE Zerjal 2019,
2020, Zhou's work).  A benefit of searching for planets in
dispersed stellar environments is that issues with stellar crowding
are also alleviated.

The identification of low-density (GIVE A NUMBER DENSITY) stellar
associations, already possible before Gaia (e.g.,
Zuckerman+ CITE), has reached a breakthrough pace with the Gaia data.
Some low-density populations (``moving groups'') have no discernable
cores (e.g., USco CITE, Psc Eri CITE, mu Tau CITE).
However other
low-density
halos seem to be centered on previously reported open clusters (e.g.,
Kounkel+19, Kounkel+20, Meingast+21).  In some
cases, these halos correspond to tidal tails, e.g., in the Hyades
(CITE Meingast+Alves 20... and Roser+20), and in the Ursa Major moving
group (CITE Gagne+20). 
However, different clustering methods on the
Gaia data tend to give different results (CITE: Hunt \& Reffert 2020).
Employing say Gaussian Mixture Modelling, or any analogous method that
requires ``clusters'' to be ellipses in phase-space (position or
velocity) unsurprisingly yields open clusters that are roughly
elliptical.  Some unsupervised clustering methods lead to comparable
``roughly elliptical'' results \citep[][hereafter
\citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
unsupervised methods such as HDBScan (e.g., \citep[][hereafter
\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}) have
been found to yield additional structures, particularly in lower
density regions such as the Psc-Eri stream (CITE: Meingast+18,
Curtis+19, Newton+21, but reword bc discovery didn't use HDBScan), but
also more generally around many open clusters
\citep{kounkel_untangling_2019}.  Unsupervised approaches that
incorporate physical constraints like imposing a maximum velocity
dispersion on putative members yield similar results, potentially with
higher purity (Meingast+21).

@@

As part of a broader project of identifying a large and clean sample of
young stars for a transit search, we focus in this paper on a 
modest question: in just a single rich southern open cluster, is the
cluster halo coeval with the core?  The cluster we chose for this
analysis was NGC\,2516, since it is young ($\sim$120\,Myr) and close
($d\approx400$\,pc) enough to facilitate rotation measurements using
TESS, and some degree of spectroscopic analysis.  We want to know: is
the halo real? To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of boring field stars?  And more
generally, what are the implications for the evolution of open clusters
if they do have halos?

@@

% TODO: this paragraph more or less parrots the processes described by
% Bouvier+20. are there others? is there a better way to describe
% these?
Both internal and external processes have been suggested to explain
the lithium-rotation trend (CITE see the recent review by Bouvier+20).
One explanation based in the stellar interior could be that the
convective mixing efficiency is anticorrelated with the surface
rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
possibility could be that stronger magnetic fields in the star's
interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
Somers + Pinsonneault 2014).  An external process that might also be
important is the effect of star-disk magnetic locking during the PMS
phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
the star's outer convective zone being locked for longer while the
radiative core contracts.  The resulting differential rotation and
rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
Eggenberger+12).
We have no preference between these possibilities---we simply note
that the \cn\ sample could be another helpful data point in
distinguishing them.

@@

\subsection{How did the halo form?}
``The stars in these halos could have escaped from the cores of the
cluster, or they could be remnants of a larger and lower-density star
formation complex that has since dispersed.''
Another possibility is {\it triggered star formation}.

@@

>>>>

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these halos could have escaped from the cores of the
  cluster, or they could be dispersed remnants of the initial star formation complex.
  They could also be field stars---the false positives of overly
  aggressive clustering algorithms.
  In this study, we focus on a halo with a length of 300\,pc reported
  around the $\approx$120\,Myr open cluster NGC\,2516.
  %TODO: says who?
  The classical tidal radius of the cluster is $\sim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two in three kinematically selected halo members 25-250\,pc from the
  cluster center show rotation periods consistent with a
  gyrochronological age of 120\,Myr.
  A comparison sample of field stars shows no such trend.
  The rotation period and lithium abundances of the cluster stars are
  also anti-correlated at fixed stellar mass, as is the case with
  comparable stellar populations.
  This work expands the set of confirmed NGC\,2516 members by a factor
  of $\approx$2, and quantifies the degree to which Gaia-based
  analyses can identify dispersed stellar populations against a
  background of field star contaminants.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

Star clusters form within hierarchically structured molecular clouds
(CITE).  Over the first $\approx$10\,Myr, ``feedback'' effects
including protostellar outflows, photoionization, radiation pressure,
and supernova shocks disperse the gas out of the cloud (CITE
Krumholz+19).  Since only a small fraction of the cloud mass is
converted into stars ($\approx 1\%$; see Krumholz+19 Figure 10), the
gas dispersal drives the majority ($\approx$90\%) of stars in the
cluster to become gravitationally unbound (e.g., Lada \& Lada 03).

From 10 to 1000\,Myr, the portion of the cluster that survives gas
dispersal suffers an onslaught that typically unbinds the cluster.
Internal to the cluster, mass loss from supernovae and AGB winds, and
close stellar encounters (``relaxation'') are important over 10s and
100s of megayears, respectively (CITE, CITE).  External to the
cluster, collisions with GMCs (Spitzer 1958), and perturbations from
the galactic tide in both the radial (CITE) and vertical (CITE)
dimensions further drive stars toward the ``high velocity tails'' of
the stellar energy distribution.

@@

Identifying the stars that have dispersed from their natal clusters
into the galactic field is of interest for a few reasons.  One 
reason is understanding the predominant conditions under which stars and star
clusters form (e.g., Wright Mamajek 2018 ScoCen), as well as
the processes that drive cluster dissolution (including gas dispersal,
internal and external heating mechanisms, and radial migration across
the galactic disk).
Another is to understand the Sun's birth conditions---
did it form in an open
cluster? If so, how massive was cluster, and is there any hope at
identifying the stars that formed near the Sun (e.g., Adams 2010)?
Finally, the dispersal of open clusters also provides perhaps
the best test case for the concept of ``chemical tagging'', aka.\
``galactic archaeology'' (CITE).  If you can't chemically tag stars
that are kinematically associated, why would you be able to do so once
they've dispersed throughout the galactic disk?

Identifying the remnant
halos\footnote{ The low-density outer regions of open clusters have
been referred to in general terms as ``halos'' and ``coronae''; we
adopt the former (CITE, CITE). } of open clusters is also important for a
separate project: that of discovering young transiting planets.  Young
transiting planets are hard to find because young stars are rare
(e.g., CITE Piskunov 2018), and reside in crowded regions of the
sky (e.g., Kharchenko 2013).  If the dispersed halos of
nearby star clusters could be reliably identified in a complete
manner, this could expand the census of nearby young stars by up to
a factor of $10$, based on the expected fraction of stars thought to
be lost during gas dispersal.

The identification of low-density (GIVE A NUMBER DENSITY) stellar
associations, though possible before Gaia (e.g.,
Zuckerman+ CITE, Bergond+01), has reached a breakthrough pace in the present era.
Some low-density moving groups have no discernable
cores (e.g., Psc Eri CITE, mu Tau CITE), or are hierarchically
sub-structured (e.g., USco CITE, Taurus[?]).
However other
low-density
halos are associated with previously known open clusters (e.g.,
Kounkel+19, Kounkel+20, Meingast+21).  In some
cases, these halos correspond to tidal tails, e.g., in the Hyades
(CITE Meingast+Alves 20, Roser+20), and in the Ursa Major moving
group (CITE Gagne+20). 
However, applying different clustering methods on the
Gaia data yields different results (CITE: Hunt \& Reffert 2020).
Using say Gaussian Mixture Modelling, or any analogous method that
requires ``clusters'' to be ellipses in phase-space unsurprisingly
yields open clusters that are roughly elliptical.  Some unsupervised
clustering methods lead to similar results \citep[][hereafter
\citetalias{cantatgaudin_gaia_2018}]{cantatgaudin_gaia_2018}.  Other
unsupervised methods
have been found to yield dispersed and asymmetric structures
with number densities down to 100 times less than the field, spread
over many hundreds of parsecs
(e.g., \citep[][hereafter
\citetalias{kounkel_untangling_2019}]{kounkel_untangling_2019}; Meingast+21).

@@

As part of a broader project of identifying a large and clean sample of
young stars for a transit search, we focus in this paper on a 
modest question: in just a single rich southern open cluster, is the
cluster halo coeval with the core?  We chose for this
NGC\,2516 for this analysis since it is young ($\sim$120\,Myr; B3V at
the turnoff) and close
($d\approx400$\,pc) enough to facilitate rotation measurements using
TESS, and some degree of spectroscopic analysis.  
Its halo, spanning $\approx$300\,pc, is also quite large.

We want to know: is
the halo real? To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of boring field stars?  And more
generally, what are the implications for the evolution of open clusters
if they do have halos?

@@

% TODO: this paragraph more or less parrots the processes described by
% Bouvier+20. are there others? is there a better way to describe
% these?
Processes both internal and external to the star have been suggested to explain
the lithium-rotation trend (CITE see the recent review by Bouvier+20).
One explanation based in the stellar interior could be that the
convective mixing efficiency is anticorrelated with the surface
rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
possibility could be that stronger magnetic fields in the star's
interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
Somers + Pinsonneault 2014).  An external process that might also be
important is the effect of star-disk magnetic locking during the PMS
phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
the star's outer convective zone being locked for longer while the
radiative core contracts.  The resulting differential rotation and
rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
Eggenberger+12).
We have no preference between these possibilities---we simply note
that the \cn\ sample could be another helpful data point in
distinguishing them.

@@

\subsection{How did the halo form?}
Another way to frame this question: should we call these stars a tidal
tail?
% TODO: read Gagne+20 on this...
The classical explanation of such a structure (e.g., Krumholz+19) is
that stars escape out of L1 and L2, and then form leading and lagging
arms due to differential rotation in the galaxy.
The leading and lagging arms of NGC\,2516's halo qualitatively
match this picture, in that they are oriented roughly toward and away
the direction of galactic rotation (CITE, Meingast21, link to their
website).

A second explanation though invokes
the idea that the cluster formed in a larger and lower-density star
formation complex, and the stars we see in the halo did not in fact
form in say the same ``clump'' as those in the cluster core.

% ``The stars in these halos could have escaped from the cores of the
% cluster, or they could be remnants of a larger and lower-density star
% formation complex that has since dispersed.''
Another possibility is {\it triggered star formation}.

@@

\subsection{Ways of doing this at different ages}
Some final discussion is warranted on the applicability of our
approach more generally.  This work starts with a kinematically
identified population, and then uses photometry (Gaia HR diagrams;
TESS rotation) and spectroscopy (GALAH and Gaia-ESO lithium) to
confirm youth in the stars.  Other approaches are also possible.

One can start with photometric criteria (e.g., Gaia HR diagram of
everything within some distance; Zari+18), or with spectroscopic
criteria (CITE Berger+18), or with combinations thereof (e.g., CITE
Zerjal 2019, 2020, Zhou's work).  
It might even be possible to start using chemical abundance signatures
(CITE, GALAHDR3?).

These paths might in certain cases succeed in identifying more
complete samples of dispersed members of young populations, since they
do not require kinematic proximity.

### 2021-02-16 planet_paper/ms.tex
update paper

\title{
  Cluster Difference Imaging Photometric Survey. IV.
  TOI 1937Ab. Young Hot Jupiter, or Just Tidally Spun Up?
}

@@

\begin{abstract}
  foo,bar.
\end{abstract}

@@

\subsection{The future}:
- Detailed multi-abundance modelling.
  i.e,. Get spectra of other cluster stars using PFS, and compare the metallicities.
- 2$\farc$5 separate, multiband imager. (With LCOGT, 2m in Australia,
  z-band)->Do PSF fitting.
- K-band image. MagAO? Multicolor imager.
- Maybe GROND -- has K-band. At La Silla. On 2.2m.  griz+K.

>>>>

\title{
  Cluster Difference Imaging Photometric Survey. IV.
  TOI 1937Ab. Youngest Hot Jupiter, or Just Tidally Spun Up?
}

@@

\begin{abstract}
  We report the discovery and confirmation of a hot Jupiter,
  TOI\,1937Ab, and present the evidence for and against its youth
  ($\approx$120\,Myr).
  We found the planet in images taken by the NASA TESS mission
  using the CDIPS pipeline.
  We measured its mass (1.X Mjup) and orbital obliquity
  (0$\pm$XX$^\circ$) using PFS at Magellan.
  Gaia kinematics suggest that the star is in the halo of NGC\,2516.
  The possible youth is corroborated by the {\bf G2V} star's 6.X day
  rotation period, the planet's 22.X hour orbital period, and the
  star's metallicity ([Fe/H] X.X$\pm$Y.Y dex) being consistent with
  that of the cluster (X.X$\pm$Z.Z dex).
  However, the star's spectrum does not show lithium, which argues
  against it being truly young.
  We outline tests that could support or refute the youth of the
  system, and outline the reasons why few, if any, sub-100 Myr hot
  Jupiters have been securely detected.
\end{abstract}

@@

\subsection{The future}:
- Detailed multi-abundance modelling.
  i.e,. Get spectra of other cluster stars using PFS, and compare the metallicities.
- 2$\farc$5 separate, multiband imager. (With LCOGT, 2m in Australia,
  z-band)->Do PSF fitting.
- K-band image. MagAO? Multicolor imager.
  E.g., Keck-NIRC2 (if it were in the North).
  ???What is the prediction for the G-K color???
- Maybe GROND -- has K-band. At La Silla. On 2.2m.  griz+K.
